---
title: "Asynch Sept 11"
date: "2023-09-11"
output: html_notebook
---
# packages
```{r}
library(tidyverse)
library(here)
library(janitor)
library(nycflights13)
```
# data

This is the data given. It contains the flights departing from NYC airports. This code chunk loads in the data.
```{r}
flights |>
  filter(dest == "IAH") |> 
  group_by(year, month, day) |> 
  summarize(
    arr_delay = mean(arr_delay, na.rm = TRUE)
  )

df1 <-  flights
```
# flights from Miami in Sept.

Below are thr flights from Miami. When it runs, there are 0 observations. This is because all the flights in the data set were departing from NYC airports.
```{r}
df2 <-  df1 |> filter(origin == "MIA", month == "9")

glimpse(df2)
```

# flights going to Miami in Sept.
By filtering the data to find destination = MIA (Miami) in September (9) there were 912 observations.
```{r}
df3 <- df1|> filter(dest == "MIA", month =="9")

glimpse(df3)
```
# flights going to Miami in Jan.

Similar to the code chunk above, destination = MIA and month = January (1).
```{r}
df4 <- df1|> filter(dest == "MIA", month =="1")

glimpse(df4)
```

# flights going to Chicago during the summer.
For "summer" months, 6,7,8 were used to capture summer. The '|' is used to represent 'or'. 
```{r}
df5 <- df1|> filter(dest == "MDW", month == (6 | 7 | 8))

glimpse(df5)
```

#flights going to Miami in September 
To sort the flight numbers and match them, the code is as follows:

We see that the smallest flight number is 83.
```{r}
df6 <- df1 |> filter (dest == "MIA", month == "9") |> slice_min(flight, n=1)

glimpse (df6)
```
# Visual gg plot

Below is the code chunk for a ggplot of the arrival delays and departure delays.
```{r}
ggplot(
  data = df6,
  mapping = aes(x = dep_delay, y = arr_delay)
)
```
Because the above code did not work and I could not find a way to correct it, I was not able to make a second visual.

# Early flights?

I found an article with similar data. It shows that when the values are negative, the flights were in fact early.
https://www.bartleby.com/questions-and-answers/through-december-7-2007.-these-data-obtained-from-bts-can-be-viewed-in-the-following-dataview-tool.-/f8678c74-815f-4d4a-89f2-9f0e66d33600

# is flight time affected by delayed departure
```{r}
df7=df1 |>
  filter(dest == "MIA", origin == "LGA") |>
  mutate(flt_delta=arr_delay-dep_delay)
```

```{r}
df7 |>
  ggplot(aes(x=dep_delay, y=flt_delta)) +geom_point(alpha=.1)
```
```{r}
df7 |>
  ggplot(aes(x=time_hour, y=dep_delay)) +
  geom_point(alpha=.1)+
  stat_smooth()+
  ylim(-25,120)
```

```{r}
df7 |>
  ggplot(aes(x=hour+minute/60, y=dep_delay)) +
  geom_point(alpha=.1)+
  stat_smooth()+
  ylim(-25,120)
```
Observation: departure delay increases as the flight day continues.

# day of the week
```{r}
df7|>
  mutate(day_of_week=weekdays(time_hour)) |>
  ggplot(aes(x=hour+minute/60, y=dep_delay, color=day_of_week)) +
  geom_point(alpha=.1) +
  stat_smooth()+
  ylim(-20,40)+
  facet_wrap(-day_of_week)
```

